---
title: "El Paso Wasterwater Surveillance Time Series Correlation Analysis"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_notebook:
    toc: yes
---

Reference: https://github.com/julianfaraway/wbequant

Three ways to deal with irregularly sampled waste water data:           
1. Interpolation - various statistical methods available           
2. Last observation carried forward: replacing each NA with the most recent non-NA prior to it.        
3. Aggregated weekly or bi-weekly.                

In this analysis, we used the second method.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message=F)
```


```{r load packages}
suppressPackageStartupMessages({
library(readr)
library(readxl)
library(dplyr)
library(stringr)
library(tidyr)
library(lubridate)
library(zoo)
library(here) #
library(svglite) #
library(tidyverse)
library(mgcv) #
library(broom)
library(jtools)
library(ggpubr)
})
```

## Data processing 

- Time window: 2020-11-1 to 2022-06-1

- Waste water measurements were sampled weekly (or bi-weekly; skipped during holidays)
  - Two sheets of data for NOV 9th 2021 (used (2))      
  - Two missing for Haskell: 1/13/22; 11/9/21     
  - Three missing for JT: 9/15/21; 9/1/21; 9/8/21 & removed data point on 10/27/22  
- Case data were collected daily.     
 
```{r read data}
# read in processed waste water data
organized.data <- read_excel("Processed_data/WeeklyReports_ElPaso_CMMR(R)_2022_Processed.xlsx") %>% mutate(Date=as.Date(`Date Collected`)) %>% 
  dplyr::rename(N1=`Copies per L waste water (N1)`, N2=`Copies per L waste water (N2)`, N1_MGD=`Copies/MGD (N1)`, N2_MGD=`Copies/MGD (N2)`) %>%
  mutate(`WWTP Site`=ifelse(`WWTP Site`=="Fred Hurvey", "Fred Hervey", `WWTP Site`))
#Start and end dates
mindate = as_date("2020-11-01")
maxdate = max(organized.data$Date, na.rm = T)
#subset the data by WWTP
FH<-subset(organized.data , `WWTP Site`=="Fred Hervey", select = c("Date", "N1", "N2", "N1_MGD", "N2_MGD")) %>% na.omit() %>% mutate(N1=N1, N2=N2) %>%
  complete(Date = seq.Date(mindate, maxdate, by="day")) %>% 
  mutate(N1 = zoo::na.locf(N1, na.rm=FALSE))

Has<-subset(organized.data , `WWTP Site`=="Haskell", select = c("Date", "N1", "N2","N1_MGD", "N2_MGD")) %>% na.omit() %>% mutate(N1=N1, N2=N2) %>%
  complete(Date = seq.Date(mindate, maxdate, by="day")) %>% 
  mutate(N1 = zoo::na.locf(N1, na.rm=FALSE))

Jth<-subset(organized.data , `WWTP Site`=="John T. Hickerson", select = c("Date", "N1", "N2","N1_MGD", "N2_MGD")) %>% na.omit() %>% mutate(N1=N1, N2=N2) %>%
  complete(Date = seq.Date(mindate, maxdate, by="day")) %>% 
  mutate(N1 = zoo::na.locf(N1, na.rm=FALSE))

Rb<-subset(organized.data , `WWTP Site`=="Roberto Bustamante", select = c("Date", "N1", "N2","N1_MGD", "N2_MGD")) %>% na.omit() %>% mutate(N1=N1, N2=N2) %>%
  complete(Date = seq.Date(mindate, maxdate, by="day")) %>% 
  mutate(N1 = zoo::na.locf(N1, na.rm=FALSE))

# read in processed case data by WWTP
case <- read_excel("Processed_data/WWTP_CaseRate_processed_updated.xlsx") %>% mutate(Date=as.Date(Date))
case <- case %>% 
  group_by(WWTP) %>% arrange(Date) %>%
  mutate(rate7avg=rollmean(rate, 7, fill=NA, align="right")) %>% ungroup() %>%
  mutate(WWTP=ifelse(WWTP=="Fred Hurvey", "Fred Hervey", WWTP))

```

## Relationship between wastewater SARS-COV2 RNA and 7-day rolling average cases per 100ks by sites

```{r fig.height=8, fig.width=5}
FH1 <- left_join(subset(case, case$WWTP=="Fred Hervey"), FH, by="Date")
Has1 <- left_join(subset(case, case$WWTP=="Haskell"), Has, by="Date")
Jth1 <- left_join(subset(case, case$WWTP=="John T. Hickerson"), Jth, by="Date")
Rb1 <-  left_join(subset(case, case$WWTP=="Roberto Bustamante"), Rb, by="Date")

par(mgp=c(1.5,0.5,0), mar=c(3.1,3.1,3.1,0), pch=20, mfrow=c(2,4))
p1 <- FH1 %>% 
  ggplot(aes(x=N1, y = rate7avg)) + 
  geom_point(size=0.5,alpha=0.25) +
  scale_x_continuous(name = "Wastewater RNA concentratiion  (copies/L)") +
  scale_y_continuous(name = "COVID-19 Case rate") +
  geom_smooth(color='black') +ggtitle("Fred Hervey - untransformed")
p2 <- FH1 %>% 
  ggplot(aes(x=N1+20, y = rate7avg)) + 
  geom_point(size=0.5,alpha=0.25) +
  scale_x_continuous(name = "Wastewater RNA concentratiion  (copies/L)", trans = "log10") +
  scale_y_continuous(name = "COVID-19 Case rate", trans = "log10") +
  geom_smooth(method="lm", color='black')+ggtitle("Fred Hervey - log transformed")
p3 <- Has1 %>%
  ggplot(aes(x=N1, y = rate7avg)) + 
  geom_point(size=0.5,alpha=0.25) +
  scale_x_continuous(name = "Wastewater RNA concentratiion  (copies/L)") +
  scale_y_continuous(name = "COVID-19 Case rate") +
  geom_smooth(color='black') +ggtitle("Haskell- untransformed")
p4 <- Has1 %>% 
  ggplot(aes(x=N1+20, y = rate7avg)) + 
  geom_point(size=0.5,alpha=0.25) +
  scale_x_continuous(name = "Wastewater RNA concentratiion  (copies/L)", trans = "log10") +
  scale_y_continuous(name = "COVID-19 Case rate", trans = "log10") +
  geom_smooth(method="lm", color='black')+ggtitle("Haskell - log transformed")
p5 <- Jth1 %>% 
  ggplot(aes(x=N1, y = rate7avg)) + 
  geom_point(size=0.5,alpha=0.25) +
  scale_x_continuous(name = "Wastewater RNA concentratiion  (copies/L)") +
  scale_y_continuous(name = "COVID-19 Case rate") +
  geom_smooth(color='black') +ggtitle("John T. Hickerson - untransformed")
p6 <- Jth1 %>%
  ggplot(aes(x=N1+20, y = rate7avg)) + 
  geom_point(size=0.5,alpha=0.25) +
  scale_x_continuous(name = "Wastewater RNA concentratiion  (copies/L)", trans = "log10") +
  scale_y_continuous(name = "COVID-19 Case rate", trans = "log10") +
  geom_smooth(method="lm", color='black')+ggtitle("John T. Hickerson - log transformed")
p7<- Rb1 %>% 
  ggplot(aes(x=N1, y = rate7avg)) + 
  geom_point(size=0.5,alpha=0.25) +
  scale_x_continuous(name = "Wastewater RNA concentratiion  (copies/L)") +
  scale_y_continuous(name = "COVID-19 Case rate") +
  geom_smooth(color='black') +ggtitle("Roberto Bustamante - untransformed")
p8 <- Rb1 %>%
  ggplot(aes(x=N1+20, y = rate7avg)) + 
  geom_point(size=0.5,alpha=0.25) +
  scale_x_continuous(name = "Wastewater RNA concentratiion  (copies/L)", trans = "log10") +
  scale_y_continuous(name = "COVID-19 Case rate", trans = "log10") +
  geom_smooth(method="lm", color='black')+ggtitle("Roberto Bustamante - log transformed")

sfigure2 <- ggarrange(p1,p2,p3,p4,p5,p6,p7,p8, ncol=2, nrow=4)

#ggsave("Sfigure2.png", sfigure2, width = 10, height=12)
sfigure2
```


## Predicting COVID-19 infection rate using concurrent wastewater measure

### Fred Hervey
```{r}
# modeling
#' Fit the a linear model with the variables on a log scale
FHmod = lm(log(rate7avg) ~ log(N1), FH1)
summ(FHmod)
```

### Haskell
```{r}
Hasmod = lm(log(rate7avg) ~ log(N1), Has1)
summ(Hasmod)
```

### John T. Hickerson
```{r}
Jthmod = lm(log(rate7avg) ~ log(N1), Jth1)
summ(Jthmod)
```

### Roberto Bustamante     
```{r}
Rbmod = lm(log(rate7avg) ~ log(N1), Rb1)
summ(Rbmod)
```



### All data

#### Correlations between N1 and incidence
```{r}
ww <- organized.data %>% select("Date", "N1", "N2", "N1_MGD", "N2_MGD", "WWTP Site") %>% na.omit() %>% mutate(N1=N1, N2=N2, WWTP=`WWTP Site`) %>%
  group_by(WWTP) %>%
  complete(Date = seq.Date(mindate, maxdate, by="day")) %>% 
  mutate(N1 = zoo::na.locf(N1, na.rm=FALSE)) %>%
  ungroup()

all <- left_join(ww, case, by=c("WWTP", "Date"))
#' Correlations between N1 and incidence
all %>% 
  group_by(WWTP) %>% 
  summarise(incidence = cor(log(N1+1),log(ncase+1),use = "complete.obs")) %>% 
  knitr::kable(digits=3)
```

#### Fit the model with all data
```{r}
#' A stronger relationship is not seen compared to the individual areas
#' 
#' Now fit a model
lmodagg = lm(log(rate7avg) ~ log(N1), all)
summ(lmodagg)
# exp(predict(lmodagg,data.frame(N1=mean(all$N1)),interval = "prediction"))
# #' compared to a FH only model
# exp(predict(FHmod,data.frame(N1=mean(all$N1)),interval = "prediction"))
```



#### Varying Coefficient models
Fit a model where each site has its own time varying intercept but the slope is in common for all:         
$$ y = \alpha_{i}(t) + \beta(t)x $$ 
```{r}
#' Assemble data
plus = all%>% 
  filter(between(Date, as.Date("2020-11-18"),as.Date("2022-06-01"))) %>% 
  mutate(y = log10(rate7avg),
         x = log10(N1 + 20),
         timet = as.numeric(Date)-18584,
         WWTP = factor(WWTP))

gmod = gam(y ~ s(timet,by=WWTP) + s(timet,by=x), data=plus)
summary(gmod)
check <- as.data.frame(gmod$coefficient)


```


```{r}
#' Plot the coefficient functions, gray lines at Jan1, Apr1 and Jul1
#+ allhabgam
sites <- unique(all$WWTP)
#par(mgp=c(1.5,0.5,0), mar=c(3.1,3.1,3.1,0), pch=20, mfrow=c(2,2))
# for(i in 1:4){
#   pgam = plot(gmod, rug=FALSE, xlab="Time", ylab='a(t)',
#        main=sites[i], select=i)
#   abline(v=c(90,180,270)+30,col=gray(0.75))
# }
# 
# #+ slopegam
# plot(gmod, rug=FALSE, xlab="Time", ylab='b(t)',
#      select=15,scale=0)
# abline(v=c(90,180,270)+30,col=gray(0.75))
#par(mfrow=c(1,1))

#' Generate nicer plots
#+ allgamplot, fig.show='hide'
#pgam = plot(gmod, pages = 1)
pgdf = data.frame(x=unlist(map(pgam,"x")),
                  fit=unlist(map(pgam,"fit")),
                  se=unlist(map(pgam,"se")),
                  WWTP = rep(c(sites,"beta"),each=100))
pgdf$ub = pgdf$fit+2*pgdf$se
pgdf$lb = pgdf$fit-2*pgdf$se
#+ pgam4
pgdf %>% filter(WWTP %in% sites[1:4]) %>% 
  ggplot(aes(x=x)) + 
  geom_ribbon(aes(ymin = lb, ymax = ub),fill="gray90") +
  geom_line(aes(y=fit)) +
  facet_wrap(~ WWTP) +
  ylab("a(t)") +
  scale_x_continuous(name="Date",breaks=c(0,44,134,224,314,404,494),
                     labels=c("11/18/20","1/1/21","4/1/21","7/1/21", "10/1/21", "1/1/22", "4/1/22"))
#+ pgamb
pgdf %>% filter(WWTP == "beta") %>% 
  ggplot(aes(x=x)) + 
  geom_ribbon(aes(ymin = lb, ymax = ub),fill="gray90") +
  geom_line(aes(y=fit)) +
  ylab("b(t)") +
  scale_x_continuous(name="Date",breaks=c(0,44,134,224,314,404,494),
                     labels=c("11/18/20","1/1/21","4/1/21","7/1/21", "10/1/21", "1/1/22", "4/1/22"))
```


## Number of WW sampling by month      
```{r}
#' Look at change in sampling over time
#+ wwtpmon
Sfigure3 <- organized.data %>% filter(Date<as.Date("2022-06-01")) %>%
  mutate(Mon = floor_date(Date, "month")) %>% 
  group_by(Mon) %>% count() %>% 
  ggplot(aes(Mon,n)) + geom_step() + geom_point()+
  xlab("Date") + ylab("Wastewater Samples per Month")+
  scale_x_date( date_breaks="month",date_labels = "%y-%m")+
  theme(axis.text.x =element_text (angle=90,hjust=1, size=14),axis.text.y =element_text (size=12), axis.title.y = element_text(size=16), axis.title.x = element_text(size=16))
Sfigure3
#ggsave("Sfigure3.png", Sfigure3, dpi=300, height=6, width=10)

# by site - panel
Sfigure3.1 <- organized.data %>% filter(Date<as.Date("2022-06-01")) %>%
  mutate(Mon = floor_date(Date, "month")) %>% 
  group_by(Mon, `WWTP Site`) %>% count() %>% 
  ggplot(aes(Mon,n)) + geom_step() + geom_point()+
  facet_wrap(~`WWTP Site`)+
  xlab("Date") + ylab("Wastewater Samples per Month")+
  scale_x_date( date_breaks="month",date_labels = "%y-%m")+
  scale_y_continuous(limits = c(0, 7), breaks = seq(0, 7, 1))+
  theme(strip.background = element_blank(),strip.text.x = element_text(size = 16),
        legend.text=element_text(size=14), axis.text=element_text(size=12), 
        axis.title = element_text(size=14),axis.text.x =element_text (angle=90,hjust=1, size=12))
Sfigure3.1
#ggsave("Sfigure3-1.png", Sfigure3.1, dpi=300, height=8, width=12)

# # by site -color NO!!!
# Sfigure3.2 <- organized.data %>% filter(Date<as.Date("2022-06-01")) %>%
#   mutate(Mon = floor_date(Date, "month")) %>% 
#   group_by(Mon, `WWTP Site`) %>% count() %>% 
#   ggplot(aes(x=Mon,y=n, color=`WWTP Site`)) + geom_step() + geom_point()+
#   xlab("Date") + ylab("Wastewater Samples per Month")+
#   scale_x_date( date_breaks="month",date_labels = "%y-%m")+
#   theme(axis.text.x =element_text (angle=90,hjust=1, size=14),axis.text.y =element_text (size=12), axis.title.y = element_text(size=16), axis.title.x = element_text(size=16))

```


## Cross-correlation analysis between reported cases and WW concentration        

Cross-correlation may help us identify the leading time when virus in WW started to show a pikes before the cases were increasing.    

The CCF is defined as the set of correlations (height of the vertical line segments in figures below between two time series $x_t + h$ (waste water concentration) and $y_t$ (case rate) for lags $h = 0, \pm1, \pm2, \ldots$. A negative value for $h$ represents a correlation between the x-series at a time before $t$ and the y-series at time $t$.               
The CCF helps to identify lags of $x_t$ that could be predictors of the $y_t$ series.           

1. When $h < 0$ , $x$ leads $y$.            
2. When $h > 0$ , $x$ lags $y$.       

### Fred Hurvey               
```{r}
FH_dt <- subset(FH1, FH1$Date <= as.Date("2022-06-01") & FH1$Date >= as.Date("2020-11-18")) %>% arrange(Date)
N1_ts <- ts(FH_dt$N1,frequency=365, start = c(2020,323))
case_ts <- ts(FH_dt$rate,frequency=365,start=c(2020,323))

## intersection (only overlapping times)
dat_int <- ts.intersect(N1_ts, case_ts)
## plot the ts
plot(dat_int, main = "Time Series Plot of WW N1 and Case Infection in Fred Hurvey", yax.flip = TRUE)

```

#### Cross-correlogram plot         
```{r}
########################
# Cross-correlogram plot
########################
## get the matching test and case data
N1_ts_match <- dat_int[,"N1_ts"]
case_ts_match <- dat_int[,"case_ts"]
## CCF of combined data
ccf <- ccf(N1_ts_match, case_ts_match, ylab = "Cross-correlation", lag.max=20, main="")
#ccf
#plot(ccf[ccf$acf>0.7])
# infection rates are relatively high ~11 days after high WW
ccf$lag[ccf$acf==max(ccf$acf)]*365

```
**Here one lag is one year, so the maximum correlation is at -11 days - Infection rates are relatively high around 11 days after high WW concentration**           


#### Scatterplots of $y_t$ versus $x_{t-h}$:
For negative lags from 0 back to a lag -15. In each plot, (WW N1 variable) is on the vertical and a past lag of infection rate is on the horizontal. Correlation values are given on each plot.         

- Correlation on the top right is the cross-correlation between WW N1 at time t and Infection rate at time t-h (h=0,1...15)        
- Red line in the plot is the lowess fit.     

```{r}
library(astsa)
lag2.plot (N1_ts_match, case_ts_match, 15)
```

<!-- Regression models -->
```{r}
# alldata=ts.intersect(case_ts,caselag1=lag(case_ts,1), caselag2=lag(case_ts_match,2), N1lag5 = lag(N1_ts_match,5),
# N1lag4=lag(N1_ts_match,4), N1lag3=lag(N1_ts_match,3), N1lag2=lag(N1_ts_match,2), N1lag1=lag(N1_ts_match,1))
# tryit = lm(case_ts_match~N1lag1 + N1lag2+N1lag3+N1lag4+N1lag5 data = alldata)
# summary (tryit)
# acf2(residuals(tryit))
```



### Haskell
```{r}
HS_dt <- subset(Has1, Has1$Date <= as.Date("2022-06-01") & Has1$Date >= as.Date("2020-11-18")) %>% arrange(Date)
N1_ts <- ts(HS_dt$N1, frequency=365, start = c(2020,323))
case_ts <- ts(HS_dt$rate, frequency=365, start = c(2020,323))

## intersection (only overlapping times)
dat_int <- ts.intersect(N1_ts, case_ts)
## plot the ts
plot(dat_int, main = "Time Series Plot of WW and Infection in Haskell", yax.flip = TRUE)

## get the matching test and case data
N1_ts_match <- dat_int[,"N1_ts"]
case_ts_match <- dat_int[,"case_ts"]
## CCF of combined data
ccf <- ccf(N1_ts_match, case_ts_match, ylab = "Cross-correlation", lag.max=20, main="")
#ccf
#plot(ccf[ccf$acf>0.7])
# infection rates are relatively high ~15 days after high WW
ccf$lag[ccf$acf==max(ccf$acf)]*365
```
**Infection rates are relatively high around 15 days after high WW concentration**         

```{r}
# scatterplot
lag2.plot (N1_ts_match, case_ts_match, 19)
```




### John T       
```{r}
JT_dt <- subset(Jth1, Jth1$Date <= as.Date("2022-06-01") & Jth1$Date >= as.Date("2020-11-18")) %>% arrange(Date)
N1_ts <- ts(JT_dt$N1, frequency=365, start = c(2020,323))
case_ts <- ts(JT_dt$rate, frequency=365, start = c(2020,323))

## intersection (only overlapping times)
dat_int <- ts.intersect(N1_ts, case_ts)
## plot the ts
plot(dat_int, main = "Time Series Plot of WW and Infection in John T.", yax.flip = TRUE)

## get the matching test and case data
N1_ts_match <- dat_int[,"N1_ts"]
case_ts_match <- dat_int[,"case_ts"]
## CCF of combined data
ccf <- ccf(N1_ts_match, case_ts_match, ylab = "Cross-correlation", lag.max=20, main="")
#ccf

#plot(ccf[ccf$acf>0.7])
# infection rates are relatively high ~8 days after high WW
ccf$lag[ccf$acf==max(ccf$acf)]*365

```
**Infection rates are relatively high around 8 days after high WW concentration**


```{r}
# scatterplot
lag2.plot (N1_ts_match, case_ts_match, 11)
```


### Roberto Bustamante

```{r}
RB_dt <- subset(Rb1, Rb1$Date <= as.Date("2022-06-01") & Rb1$Date >= as.Date("2020-11-18")) %>% arrange(Date)
N1_ts <- ts(RB_dt$N1, frequency=365, start = c(2020,323))
case_ts <- ts(RB_dt$rate, frequency=365, start = c(2020,323))

## intersection (only overlapping times)
dat_int <- ts.intersect(N1_ts, case_ts)
## plot the ts
plot(dat_int, main = "Time Series Plot of WW and Infection in Roberto Bustamante", yax.flip = TRUE)

## get the matching test and case data
N1_ts_match <- dat_int[,"N1_ts"]
case_ts_match <- dat_int[,"case_ts"]
## CCF of combined data
ccf <- ccf(N1_ts_match, case_ts_match, ylab = "Cross-correlation", lag.max=20, main="")
#ccf
#plot(ccf[ccf$acf>0.7])
# infection rates are relatively high ~8 days after high WW
ccf$lag[ccf$acf==max(ccf$acf)]*365
```
**Infection rates are relatively high around 8 days after high WW concentration**


```{r}

# scatterplot
lag2.plot (N1_ts_match, case_ts_match, 11)
```




## Lagged Models for Prediction       
```{r}
#' 
#' Can we predict the future? Try lagging the N1 variable:
#' 
timewindow = 30
ldf = tibble(lag = 0:timewindow, r2 = rep(NA, timewindow+1),
             se = rep(NA,timewindow+1))

for(i in 0:timewindow){
  all %>% mutate(N1lag = lag(N1,i)) -> lag
  lmodlag = lm(log(rate7avg) ~ log(N1lag+20), lag)
  gd = glance(lmodlag)
  ldf$r2[i+1] = gd$r.squared
  ldf$se[i+1] = gd$sigma
}

ldf #%>% knitr::kable(digits = 3)
ggplot(ldf) + geom_line(aes(x=lag, y=r2)) + geom_vline(xintercept=7, color="red")
```

```{r}
#' Consider trying to predict a week ahead:

all %>% mutate(rollperlag = lag(rate7avg,7)) -> lag
lmodlag7 = lm(log(rate7avg) ~ log(rollperlag), lag)
glance(lmodlag7)[c('r.squared','sigma')]

```

```{r}
#' This model predicts a week ahead using current cases. We can
#' see this works better than using N1. Now suppose
#' we have the current N1 count and we add this information to the model:
#' 
#' 
all %>% #group_by(WWTP) %>% 
  mutate(rollperlag = lag(rate7avg,7),
         N1lag7 = lag(N1,7)) -> lag
lmodlagcase7 = lm(log(rate7avg) ~ (log(rollperlag) + log(N1lag7+20)), lag)
glance(lmodlagcase7)[c('r.squared','sigma')]
```

## Prediction by site

### Fred Hervey
```{r}

timewindow = 30
ldf = tibble(lag = 0:timewindow, r2 = rep(NA, timewindow+1),
             se = rep(NA,timewindow+1))

for(i in 0:timewindow){
  FH1 %>% mutate(N1lag = lag(N1,i)) -> lag
  lmodlag = lm(log(rate7avg) ~ log(N1lag+20), lag)
  gd = glance(lmodlag)
  ldf$r2[i+1] = gd$r.squared
  ldf$se[i+1] = gd$sigma
}
ldf1 <- ldf %>% mutate(WWTP="Fred Hervey",Max=ifelse(r2==max(r2), "Yes", "No")) #%>% knitr::kable(digits = 3)


g1 <- ggplot(ldf) + 
  geom_line(aes(x=lag, y=r2, colour="R2"))+ 
  geom_line(aes(x=lag, y=se, colour="SE"), data=ldf)+
  #geom_point(aes(x=lag, y=se/2, color="blue"))+geom_line(color="blue")+
  #scale_color_manual(name = NULL, values = c("R2" = "red", "SE"="blue")) +
  labs(x="Lag (days)", y=expression(R^2), color="")+
  scale_x_continuous(breaks=c(0,5,10,15,20,25,30))+
  ggtitle("Fred Hervey")
  # scale_y_continuous(   # Custom the Y scales:
  #   name = "R2", # Features of the first axis
  #   sec.axis = sec_axis(trans=~.*1.25, name="SE"))+ # Add a second axis and specify its features

FH1 %>% mutate(rollperlag = lag(rate7avg,7)) -> lag
lmodlag7 = lm(log(rate7avg) ~ log(rollperlag), lag)
glance(lmodlag7)[c('r.squared','sigma')]

FH1 %>% #group_by(WWTP) %>% 
  mutate(rollperlag = lag(rate7avg,7),
         N1lag7 = lag(N1,7)) -> lag
lmodlagcase7 = lm(log(rate7avg) ~ (log(rollperlag) + log(N1lag7+20)), lag)
glance(lmodlagcase7)[c('r.squared','sigma')]
```

### Haskell
```{r}
timewindow = 30
ldf = tibble(lag = 0:timewindow, r2 = rep(NA, timewindow+1),
             se = rep(NA,timewindow+1))

for(i in 0:timewindow){
  Has1 %>% mutate(N1lag = lag(N1,i)) -> lag
  lmodlag = lm(log(rate7avg) ~ log(N1lag+20), lag)
  gd = glance(lmodlag)
  ldf$r2[i+1] = gd$r.squared
  ldf$se[i+1] = gd$sigma
}
ldf2 <- ldf %>% mutate(WWTP="Haskell",Max=ifelse(r2==max(r2), "Yes", "No"))#%>% knitr::kable(digits = 3)
g2 <- ggplot(ldf) + 
  geom_line(aes(x=lag, y=r2, colour="R2"))+ 
  geom_line(aes(x=lag, y=se, colour="SE"), data=ldf)+
  #geom_point(aes(x=lag, y=se/2, color="blue"))+geom_line(color="blue")+
  #scale_color_manual(name = NULL, values = c("R2" = "red", "SE"="blue")) +
  labs(x="Lag (days)", y=expression(R^2), color="")+
  scale_x_continuous(breaks=c(0,5,10,15,20,25,30))+
  ggtitle("Haskell")

Has1 %>% mutate(rollperlag = lag(rate7avg,7)) -> lag
lmodlag7 = lm(log(rate7avg) ~ log(rollperlag), lag)
glance(lmodlag7)[c('r.squared','sigma')]

Has1 %>% #group_by(WWTP) %>% 
  mutate(rollperlag = lag(rate7avg,7),
         N1lag7 = lag(N1,7)) -> lag
lmodlagcase7 = lm(log(rate7avg) ~ (log(rollperlag) + log(N1lag7+20)), lag)
glance(lmodlagcase7)[c('r.squared','sigma')]
```

### John T.
```{r}
timewindow = 30
ldf = tibble(lag = 0:timewindow, r2 = rep(NA, timewindow+1),
             se = rep(NA,timewindow+1))

for(i in 0:timewindow){
  Jth1 %>% mutate(N1lag = lag(N1,i)) -> lag
  lmodlag = lm(log(rate7avg) ~ log(N1lag+20), lag)
  gd = glance(lmodlag)
  ldf$r2[i+1] = gd$r.squared
  ldf$se[i+1] = gd$sigma
}
ldf3 <- ldf %>% mutate(WWTP="John T. Hickerson", Max=ifelse(r2==max(r2), "Yes", "No"))#%>% knitr::kable(digits = 3)
g3 <- ggplot(ldf) + 
  geom_line(aes(x=lag, y=r2, colour="R2"))+ 
  geom_line(aes(x=lag, y=se, colour="SE"), data=ldf)+
  #geom_point(aes(x=lag, y=se/2, color="blue"))+geom_line(color="blue")+
  #scale_color_manual(name = NULL, values = c("R2" = "red", "SE"="blue")) +
  labs(x="Lag (days)", y=expression(R^2), color="")+
  scale_x_continuous(breaks=c(0,5,10,15,20,25,30))+
  ggtitle("John T. Hickerson")

Jth1 %>% mutate(rollperlag = lag(rate7avg,7)) -> lag
lmodlag7 = lm(log(rate7avg) ~ log(rollperlag), lag)
glance(lmodlag7)[c('r.squared','sigma')]

Jth1 %>% #group_by(WWTP) %>% 
  mutate(rollperlag = lag(rate7avg,7),
         N1lag7 = lag(N1,7)) -> lag
lmodlagcase7 = lm(log(rate7avg) ~ (log(rollperlag) + log(N1lag7+20)), lag)
glance(lmodlagcase7)[c('r.squared','sigma')]
```

### Roberto Bustamante 
```{r}
timewindow = 30
ldf = tibble(lag = 0:timewindow, r2 = rep(NA, timewindow+1),
             se = rep(NA,timewindow+1))

for(i in 0:timewindow){
  Rb1 %>% mutate(N1lag = lag(N1,i)) -> lag
  lmodlag = lm(log(rate7avg) ~ log(N1lag+20), lag)
  gd = glance(lmodlag)
  ldf$r2[i+1] = gd$r.squared
  ldf$se[i+1] = gd$sigma
}
ldf4 <- ldf %>% mutate(WWTP="Roberto Bustamante",Max=ifelse(r2==max(r2), "Yes", "No")) #%>% knitr::kable(digits = 3)
g4 <- ggplot(ldf) + 
  geom_line(aes(x=lag, y=r2, colour="R2"))+ 
  geom_line(aes(x=lag, y=se, colour="SE"), data=ldf)+
  #geom_point(aes(x=lag, y=se/2, color="blue"))+geom_line(color="blue")+
  #scale_color_manual(name = NULL, values = c("R2" = "red", "SE"="blue")) +
  labs(x="Lag (days)", y=expression(R^2), color="")+
  scale_x_continuous(breaks=c(0,5,10,15,20,25,30))+
  ggtitle("Roberto Bustamante")

Rb1 %>% mutate(rollperlag = lag(rate7avg,7)) -> lag
lmodlag7 = lm(log(rate7avg) ~ log(rollperlag), lag)
glance(lmodlag7)[c('r.squared','sigma')]

Rb1 %>% #group_by(WWTP) %>% 
  mutate(rollperlag = lag(rate7avg,7),
         N1lag7 = lag(N1,7)) -> lag
lmodlagcase7 = lm(log(rate7avg) ~ (log(rollperlag) + log(N1lag7+20)), lag)
glance(lmodlagcase7)[c('r.squared','sigma')]
```

```{r}
ldf_all <- bind_rows(ldf1, ldf2, ldf3, ldf4)
Figure3 <-ggplot(ldf_all) + 
  geom_line(aes(x=lag, y=r2, colour="R2"))+ 
  geom_point(aes(x=lag, y=r2),color="Red", data=subset(ldf_all, ldf_all$Max=="Yes"))+
  geom_line(aes(x=lag, y=se, colour="SE"))+
  facet_wrap(~WWTP, nrow=2, ncol=2)+
  #geom_point(aes(x=lag, y=se/2, color="blue"))+geom_line(color="blue")+
  #scale_color_manual(name = NULL, values = c("R2" = "red", "SE"="blue")) +
  labs(x="Lag (days)", y=expression(R^2), color="")+
  scale_x_continuous(breaks=c(0,5,10,15,20,25,30))+
  theme_bw()+
  theme(strip.background = element_blank(),strip.text.x = element_text(size = 16),
        legend.text=element_text(size=14), axis.text=element_text(size=12), 
        axis.title = element_text(size=14))
Figure3
#ggsave("Figure3.png", Figure3, dpi=300, width=10, height=7)
```


